---
title: "Alessandra Gudoy - R Project 3"
author: "Alessandra Gudoy"
date: "10/12/2023"
output: html_document
---

```{r Libraries, echo=TRUE, warning=FALSE, messsage=FALSE}
library(ggplot2)
library(tidyverse)
```

- ### Reading categorical data
```{r Data, echo=TRUE, messsage=FALSE, warning=FALSE}

categories = c("A+","A+","A+","A+","A+","A+","A","A","A","A","A","A","A","A","A","A-","A-","A-","A-","A-","A-","A-","A-","A-","A-","A-","A-","A-","A-","B+","B+","B+","B+","B+","B+","B+","B+","B","B","B","B","B-","B-","B-","B-","B-","B-","C+","C+","C","C","C","C","C","C","C-","D+","D","F","F")

counts = table(categories)
counts

```

- ### Bar plots for categorical data
```{r Barplots, echo=TRUE, messsage=FALSE, warning=FALSE}

texture = c(5, 15, 30, 5, 15, 30, 5, 15, 30, 5, 15, 30)

barplot(counts, main = "Grades", ylab="Frequency", density=texture, col="blue")

barplot(counts/length(categories), main = "Grades", ylab="Relative Frequency", density=texture, col="red")

```

- ### Simple pie chart for categorical data
```{r Piechart, echo=TRUE, messsage=FALSE, warning=FALSE}

pie(counts, col=c(1:12), main = "Student Letter Grade")

```

- ### Fancy pie chart using "gglot2" package
```{r ggplot2 Piechart, echo=TRUE, warning=FALSE, messsage=FALSE}

df = as.data.frame(counts)

ggplot(df, aes(x="", y=Freq, fill=categories)) + 
  geom_bar(width=1, stat="identity") + 
  coord_polar("y", start=0) +
      theme_void() + theme(legend.position="none") + 
      geom_text(aes(label=categories), position = position_stack(vjust = 0.5), color = "white", size=2)

```

- ### Reading numerical data into a dataframe
```{r Student Data Frame, ggplot2 Piechart, echo=TRUE, warning=FALSE, messsage=FALSE}

Student = data.frame(
  id = c(1:60), 
  
  HW = c(96.93, 80.84, 99.62, 100, 90.8, 92.72, 80.08, 98.08, 96.55, 92.34, 44.06, 98.47, 95.79, 100,
         93.87, 91.19, 25.67, 71.65, 98.85, 100, 100, 100, 90.04, 91.57, 100, 96.17, 90.42, 91.57, 98.47,
         100, 100, 96.17, 68.58, 99.23, 100, 99.23, 97.7, 100, 95.02, 88.51, 98.85, 91.57, 83.14, 93.1,
         10.73, 100, 93.49, 86.21, 88.12, 99.62, 100, 100, 99.62, 45.98, 91.19, 97.32, 98.08, 98.08,
         99.23, 97.7), 
  
  XM = c(87.5, 42.08, 97.91, 91.66, 70.83, 79.16, 81.25, 77.08, 87.5, 68.75, 75, 87.5, 83.33, 95.83,
         85.41, 60.41, 68.75, 62.5, 81.25, 83.33, 89.58, 95.83, 77.08, 79.16, 83.33, 77.08, 32.08, 79.16,
         93.75, 97.91, 91.66, 77.08, 70.83, 87.5, 72.91, 85.41, 85.41, 75, 91.66, 79.16, 87.5, 87.5,
         85.41, 70.83, 32.08, 95.83, 79.16, 93.75, 77.08, 10.41, 93.75, 95.83, 95.83, 32.08, 87.5, 93.75,
         79.16, 83.33, 95.83, 100), 
  
  QZ = c(72.73, 40, 89.09, 78.18, 56.37, 63.64, 56.37, 56.37, 69.09, 36.37, 47.27, 63.64, 76.37, 90.91,
         60, 52.73, 25.46, 45.46, 52.73, 78.18, 63.64, 90.91, 43.64, 78.18, 67.27, 54.55, 47.27, 47.27,
         74.55, 69.09, 94.55, 36.37, 56.37, 60, 80, 83.64, 78.18, 76.37, 61.82, 65.46, 85.46, 63.64,
         74.55, 41.82, 3.64, 78.18, 63.64, 72.73, 50.91, 38.18, 100, 89.09, 80, 76.37, 52.73, 78.18,
         54.55, 63.64, 76.37, 78.18) )

```

- ### Descriptive Statistics
```{r Descriptive Statistics, echo=TRUE, warning=FALSE, message=FALSE}

summary(Student[,2:4])

```

- ### Boxplots
```{r Boxplot, echo=TRUE, warning=FALSE, message=FALSE}

boxplot(Student[,2:4], main="Student Grades", xlab="Assignment", ylab="Grade", border="palevioletred1", col="lightpink1")

```

- ### Histogram for exam grades
```{r Exam Grade Histograms, echo=TRUE, warning=FALSE, message=FALSE}

hist(Student$XM, 
 col="lightskyblue", 
 border="black",
 prob = TRUE, 
 xlab = "Exam grades",
 ylim = c(0, 0.035),
 main = "Exam Grades")
 lines(density(Student$XM), # this adds density line
 lwd = 2, # thickness of line
 col = "blue")

```

- ### Correlation matrix for numerical data
```{r Correlation Matrix, echo=TRUE, warning=FALSE, message=FALSE}

round(cor(Student[,2:4]), 2)

```

- ### Scatterplot for numerical data
```{r Scatterplot, echo=TRUE, warning=FALSE, message=FALSE}

plot(Student$HW, Student$XM, pch=20, main = "Homework & Quiz Grades vs Exam Grade", xlab = "Homework (black) and Quiz (red)", ylab = "Exam Grade")
points(Student$QZ, Student$XM, pch=20, col= "red")

```

- ### Adding a least square line to the scatter plot
```{r LSRL, echo=TRUE, warning=FALSE, message=FALSE}

plot(Student$HW, Student$XM, pch=20, main = "Homework & Quiz Grades vs Exam Grade", xlab = "Homework (black) and Quiz (red)", ylab = "Exam Grade")
points(Student$QZ, Student$XM, pch=20, col= "red")

abline(lm(Student$XM ~ Student$HW)) # lm() is the code for OLS least Square Line
abline(lm(Student$XM ~ Student$QZ), col="red")

```

- #### What are the slope and the intercept of the Least Square Line for y = XM and x = HW?
```{r echo=TRUE, warning=FALSE, message=FALSE}
s=summary(lm(Student$XM ~ Student$HW))
```
      - The slope is b = `r s$coefficients[2]` and the intercept is a = `r s$coefficients[1]`


- #### What are the slope and the intercept of the Least Square Line for y = XM and x = QZ?
```{r echo=TRUE, warning=FALSE, message=FALSE}
s=summary(lm(Student$XM ~ Student$QZ))
```
      - The slope is b = `r s$coefficients[2]` and the intercept is a = `r s$coefficients[1]`
