---
title: "Alessandra Gudoy - R Project 4"
author: "Alessandra Gudoy"
date: "12/08/2023"
output: html_document
---

```{r Libraries, echo=TRUE, warning=FALSE, messsage=FALSE}
library(ggplot2)
library(palmerpenguins)
library(car)
library(ggstatsplot)
```

```{r Getting Data, echo=TRUE, warning=FALSE, messsage=FALSE}

dat <- data.frame(species = penguins$species, flipper_length_mm = penguins$flipper_length_mm)

summary(dat)

```

The function summary(data) outputs the summary statistics of the given data (the flipper length of three different species of penguins-Adelie, Chinstrap, Gentoo). It outputs the five-number summary (the minimum length, the first quartile, median, third quartile, and maximum length), the mean of the flipper lengths, and the number of values missing, and therefore, excluded (NA, last row of table).

From the data, we can see that 152 flipper lengths came from the Adelie species, 68 from the Chinstrap species, and 124 from the Gentoo species. Two values are missing (one Adelie, one Gentoo). The minimum flipper length is 172.0 mm, the maximum flipper length is 231.0 mm, the median is 197.0 mm. The lower (or first) quartile is 190.0 mm and the upper (or third) quartile is 213.0 mm. The mean of the flipper lengths is 200.9. Because the median and mean are close in value, we can guess that the distribution is more or less symmetric.

```{r PLOTTING, echo=TRUE, warning=FALSE, messsage=FALSE}

# SCATTERPLOT
ggplot(dat) +
  aes(x = species, y = flipper_length_mm, color = species) +
  geom_jitter() +
  theme(legend.position = "none")
```

A plot showing the different flipper lengths of the different species, where the species are on the x-axis, while the length of the flippers (in mm) are on the y-axis. The red dots represent the flipper lengths for the Adelie species, green for Chinstrap, and blue for Gentoo.

The plot shows that the Gentoo may have a flippers that are a bit longer than the Adelie penguins, while the Gentoo species have longer flippers than both Gentoo penguins and Adelie penguins.

```{r BOXPLOT, echo=TRUE, warning=FALSE, message=FALSE}
# BOXPLOT
boxplot(flipper_length_mm ~ species, data = dat)
```

A boxplot of the flipper lengths of the different species, where each section of each boxplot contains 25% of the data, the darkest line is the median, the box contains 50% of the data, and circles are outliers determined by the 1.5xIQR Rule.

There are two outliers for the Adelie penguins, while there are no outliers for the Chinstrap and Gentoo penguins. All three distributions seem to be symmetric. The Adelie penguins have the lowest median for flipper length, while the Gentoo distribution has the highest median. The Gentoo distribution seems to has the lowest amount of variance/spread, while the Adelie distribution has the greatest amount of variance/spread.

```{r HISTOGRAM & NORMALITY PLOT, echo=TRUE, warning=FALSE, messsage=FALSE}
res_anova <- aov(flipper_length_mm ~ species, data = dat)

par(mfrow = c(1, 2)) # combine plots

# Histogram
hist(res_anova$residuals)

# Normality plot
qqPlot(res_anova$residuals,  id = FALSE ) # id = FALSE to remove point identification
```

```{r EQUAL VARIANCES, echo=TRUE, warning=FALSE, messsage=FALSE}

# 1.Equal variances
plot(res_anova, which = 3)

```

```{r BOX/VIOLIN PLOTS, echo=TRUE, warning=FALSE, messsage=FALSE}

ggbetweenstats(
  data = dat,
  x = species,
  y = flipper_length_mm,
  type = "parametric", # ANOVA or Kruskal-Wallis
  var.equal = TRUE, # ANOVA or Welch ANOVA
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = FALSE,
  bf.message = FALSE
)

```

```{r SUMMARY OF ANOVA, echo=TRUE, warning=FALSE, messsage=FALSE}

summary(res_anova)

```
 

