---
title: "Alessandra Gudoy - R Project 4"
author: "Alessandra Gudoy"
date: "12/08/2023"
output: html_document
---

```{r Libraries, echo=TRUE, warning=FALSE, messsage=FALSE}
library(ggplot2)
library(palmerpenguins)
library(car)
library(ggstatsplot)
```

```{r Getting Data, echo=TRUE, warning=FALSE, messsage=FALSE}

dat <- data.frame(species = penguins$species, flipper_length_mm = penguins$flipper_length_mm)

summary(dat)

```

The function summary(data) outputs the summary statistics of the given data (the flipper length of three different species of penguins-Adelie, Chinstrap, Gentoo). It outputs the five-number summary (the minimum length, the first quartile, median, third quartile, and maximum length), the mean of the flipper lengths, and the number of values missing, and therefore, excluded (NA, last row of table).

```{r PLOTTING, echo=TRUE, warning=FALSE, messsage=FALSE}

# SCATTERPLOT
ggplot(dat) +
  aes(x = species, y = flipper_length_mm, color = species) +
  geom_jitter() +
  theme(legend.position = "none")
```

A plot showing the different flipper lengths of the different species, where the species are on the x-axis, while the length of the flippers (in mm) are on the y-axis.

```{r BOXPLOT, echo=TRUE, warning=FALSE, message=FALSE}
# BOXPLOT
boxplot(flipper_length_mm ~ species, data = dat)
```

A boxplot of the flipper lengths of the different species, where each section of each boxplot contains 25% of the data, the darkest line is the median, the box contains 50% of the data, and circles are outliers determined by the 1.5xIQR Rule.

```{r HISTOGRAM & NORMALITY PLOT, echo=TRUE, warning=FALSE, messsage=FALSE}
res_anova <- aov(flipper_length_mm ~ species, data = dat)

par(mfrow = c(1, 2)) # combine plots

# Histogram
hist(res_anova$residuals)

# Normality plot
qqPlot(res_anova$residuals,  id = FALSE ) # id = FALSE to remove point identification
```

```{r EQUAL VARIANCES, echo=TRUE, warning=FALSE, messsage=FALSE}

# 1.Equal variances
plot(res_anova, which = 3)

```

```{r BOX/VIOLIN PLOTS, echo=TRUE, warning=FALSE, messsage=FALSE}

ggbetweenstats(
  data = dat,
  x = species,
  y = flipper_length_mm,
  type = "parametric", # ANOVA or Kruskal-Wallis
  var.equal = TRUE, # ANOVA or Welch ANOVA
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = FALSE,
  bf.message = FALSE
)

```

```{r SUMMARY OF ANOVA, echo=TRUE, warning=FALSE, messsage=FALSE}

summary(res_anova)

```


